Intentional release of native species undermines ecological stability

### 2022-10-19 rmd/manuscript.Rmd
Update manuscript.Rmd

Human demands for natural resources are ever-increasing, such that active interventions are critical to the sustainable management of fisheries, forestry, and wildlife [@laikreCompromisingGeneticDiversity2010]. Releases of captive-bred native species ("intentional release") is a form of the efforts to enhance wild populations of diverse taxa [@snyderLimitationsCaptiveBreeding1996; @laikreCompromisingGeneticDiversity2010]. In particular, the magnitude of intentional release for commercial and recreational species is massive (reviewed in [@laikreCompromisingGeneticDiversity2010; @kitadaEconomicEcologicalGenetic2018]) -- a single program releases up to tens of millions of plant [@laikreReleaseAlienPopulations2006], invertebrate (bivalves, insects) [@ingsBumblebeesHumblePollinators2005; @haagNorthAmericanFreshwater2012], and vertebrate individuals (fish, birds) [@negroRAPDAnalysisDetection2001; @kitadaEconomicEcologicalGenetic2018] into the wild annually. There is growing awareness that intentional release entails ecological risks, including the accumulation of deleterious alleles [@arakiGeneticEffectsCaptive2007], the spread of diseases in wild populations [@krkosekDecliningWildSalmon2007], and the intensification of intraspecific competition within the enhanced species [@kitadaLessonsLearnedJapanese2006]. Yet, this method is still pervasive in natural resource management owing to the significant economic benefit [@malkamakiSystematicReviewSocioeconomic2018; @davidEconomicFeasibilityIntensification2018; @rutledgeCostBenefitHatcheryreared1990; @kitadaEconomicEcologicalGenetic2018]. For example, Kitada @kitadaEconomicEcologicalGenetic2018 analyzed the economic performance of release programs for 12 major fishery resources across the globe. The gross monetary yield exceeded the seed-production cost in nine species, suggesting that release programs are economically profitable in general.

>>>>

Human demands for natural resources are ever-increasing, such that active interventions are critical to the sustainable management of fisheries, forestry, and wildlife [@laikreCompromisingGeneticDiversity2010]. Releases of captive-bred native species ("intentional release") is a form of the efforts to enhance wild populations of diverse taxa [@snyderLimitationsCaptiveBreeding1996; @laikreCompromisingGeneticDiversity2010]. In particular, the magnitude of intentional release for commercial and recreational species is massive (reviewed in [@laikreCompromisingGeneticDiversity2010; @kitadaEconomicEcologicalGenetic2018]) -- a single program releases up to tens of millions of plant [@laikreReleaseAlienPopulations2006], invertebrate (bivalves, insects) [@ingsBumblebeesHumblePollinators2005; @haagNorthAmericanFreshwater2012], and vertebrate individuals (fish, birds) [@negroRAPDAnalysisDetection2001; @kitadaEconomicEcologicalGenetic2018] into the wild annually. There is growing awareness that intentional release entails ecological risks, such as the accumulation of deleterious alleles and the intensified competition within the [@arakiGeneticEffectsCaptive2007; @krkosekDecliningWildSalmon2007; @kitadaLessonsLearnedJapanese2006]. Yet, this method is still pervasive in natural resource management owing to the significant economic benefit [@malkamakiSystematicReviewSocioeconomic2018; @davidEconomicFeasibilityIntensification2018; @rutledgeCostBenefitHatcheryreared1990; @kitadaEconomicEcologicalGenetic2018]. For example, Kitada @kitadaEconomicEcologicalGenetic2018 analyzed the economic performance of release programs for 12 major fishery resources across the globe. The gross monetary yield exceeded the seed-production cost in nine species, suggesting that release programs are economically profitable in general.
